---
title: "MetaMap example code"
author: "Lukas Simon"
date: "4th May 2018"
output: html_document
---

This R code will reduce the MetaMap data to project SRP066090 (Zheng et al.) and run differential metafeature abundance analysis reproducing parts of Figure 2 in the manuscript.

Load R libraries
```{r warning=FALSE}
suppressMessages(library(Matrix))
suppressMessages(library(DESeq2))
suppressMessages(library(ggrepel))
suppressMessages(library(ggplot2))
```

Load the data
```{r}
load("MetaMapData.RData")
```

Restrict data to study SRP066090 
```{r}
ok <- which(sample_info$study == 'SRP066090')
sample_info <- sample_info[ok,]
meta_count <- meta_count[,ok]
```

Identify treatment
```{r}
treat <- rep("positive", ncol(meta_count))
treat[grep("Neg", sample_info$sample_attribute)] <- "negative"
```

Run differential metafeature abundance analysis
```{r}
des <- DESeqDataSetFromMatrix(countData = round(meta_count), colData = data.frame(treat = treat), design = ~ treat)
sizeFactors(des) <- log10(sample_info$spots) / median(log10(sample_info$spots))
res <- results(DESeq(des), contrast = c('treat', 'positive', 'negative'))
```

Generate differential expression volcano plot
```{r}
tmp <- data.frame(res)
tmp <- tmp[sort.list(tmp$pvalue),]
tmp <- tmp[which(!is.na(tmp$log2FoldChange)),]
speciesNames.zhang <- meta_info[match(rownames(tmp), meta_info[,'TaxID']), 'Species']
makeVolcano <- function(res, nom){
  ggplot(tmp, aes(log2FoldChange, -log10(pvalue))) +
  geom_point(color = rgb(0,0,0,0.5)) +
  geom_point(data = tmp[1, ], aes(x=log2FoldChange, y=-log10(pvalue)), color="red") +
  ggtitle(nom) +
  theme_classic(base_size = 16) +
  labs(x = "Fold change (log2)", y = "-log10 p-value")
}
makeVolcano(res, nom = 'Zhang et al')
```

Plot Alphapapilloma levels by treatment
```{r}
abundance <- t(t(meta_count + 1) / sample_info$spots) * 1e6
species <- abundance[which(rownames(abundance) == "337041"),] 
aframe <- data.frame(Alphapapillomavirus.9 = species, treat)
ggplot(aframe, aes(treat, Alphapapillomavirus.9)) + geom_boxplot() + geom_jitter(width = 0.2) + theme(axis.title.x=element_blank()) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
